Interpreting Random Forest Predictions for Firearm Identification Using LIME

### 2018-12-04 paper/Interpreting Random Forest Predictions for Firearm Identification Using LIME.Rmd
worked on visualizing and summarizing sensativity analysis results

```{r}
# Create a subsampled version of the training data
if (!file.exists("../data/hamby173and252_train_sub.csv")) {

  # Create the data (with a set seed)
  set.seed(1837592)
  hamby173and252_train_sub <- hamby173and252_train %>%
    filter(samesource == FALSE) %>%
    slice(sample(1:length(hamby173and252_train$barrel1), 4500, replace = FALSE)) %>%
    bind_rows(hamby173and252_train %>% filter(samesource == TRUE))

  # Save the subsampled data
  write.csv(hamby173and252_train_sub, "../data/hamby173and252_train_sub.csv", row.names = FALSE)

  # Load in the subsampled training data
  hamby173and252_train_sub <- read.csv("../data/hamby173and252_train_sub.csv")

@@

```{r}
hamby224_lime_results <- hamby224_lime_comparisons %>%
  filter(!is.na(rfscore)) %>%
  group_by(training_data, bin_continuous, quantile_bins, nbins, use_density, set) %>%
  summarise(mse = (sum(diff)^2) / length(diff),
            ave_r2 = mean(model_r2)) %>%
  arrange(training_data, set)
```

```{r}
hamby224_lime_results %>%
  group_by(training_data, set) %>%
  arrange(mse) %>%
  slice(1:2)
```

```{r}
hamby224_lime_results %>%
  ungroup() %>%
  mutate(bc_qb = as.character(factor(bin_continuous):factor(quantile_bins)),
         nbins = factor(ifelse(bin_continuous == FALSE, NA, as.character(nbins)))) %>%
  ggplot(aes(x = nbins, y = mse, color = bc_qb, group = bc_qb)) +
  geom_point() + 
  geom_line() + 
  facet_grid(training_data ~ set) + 
  theme_bw()
```

```{r}
hamby224_lime_results %>%
  group_by(training_data, set) %>%
  arrange(desc(ave_r2)) %>%
  slice(1:2)
```

```{r}
hamby224_lime_results %>%
  ungroup() %>%
  mutate(bc_qb = as.character(factor(bin_continuous):factor(quantile_bins)),
         nbins = factor(ifelse(bin_continuous == FALSE, NA, as.character(nbins)))) %>%
  ggplot(aes(x = nbins, y = ave_r2, color = bc_qb, group = bc_qb)) +
  geom_point() + 
  geom_line() + 
  facet_grid(training_data ~ set) + 
  theme_bw()
```

@@

@@

Ideas for improvement of LIME: 
  - change the binning method:
    - lime is not good with linear relationships with classifiers due to the 
    inside-outside binning - it would be better to use a cumulative approach
  - consider interactions in the ridge regression model
  - learn lambda in the ridge regression
  - could add a penalty for the number of parameters in the model (or the number of bins)
  - could start with a fine grid of bins and then go backwards and fit the models on this sampled data
  - could try a type of ANOVA test if we can have nested models or if we can assume nesting
  - could compute something like an R^2_adj

>>>>

@@

```{r}
hamby224_lime_results <- hamby224_lime_comparisons %>%
  filter(!is.na(rfscore)) %>%
  group_by(bin_continuous, quantile_bins, nbins, use_density, set) %>%
  summarise(mse = (sum(diff)^2) / length(diff),
            ave_r2 = mean(model_r2)) %>%
  arrange(set)
```

```{r}
hamby224_lime_results %>%
  group_by(set) %>%
  arrange(mse) %>%
  slice(1:2)
```

```{r}
hamby224_lime_results %>%
  ungroup() %>%
  mutate(bc_qb = as.character(factor(bin_continuous):factor(quantile_bins)),
         nbins = factor(ifelse(bin_continuous == FALSE, NA, as.character(nbins)))) %>%
  ggplot(aes(x = nbins, y = mse, color = bc_qb, group = bc_qb)) +
  geom_point() + 
  geom_line() + 
  facet_grid( ~ set) + 
  theme_bw()
```

```{r}
hamby224_lime_results %>%
  group_by(set) %>%
  arrange(desc(ave_r2)) %>%
  slice(1:2)
```

```{r}
hamby224_lime_results %>%
  ungroup() %>%
  mutate(bc_qb = as.character(factor(bin_continuous):factor(quantile_bins)),
         nbins = factor(ifelse(bin_continuous == FALSE, NA, as.character(nbins)))) %>%
  ggplot(aes(x = nbins, y = ave_r2, color = bc_qb, group = bc_qb)) +
  geom_point() + 
  geom_line() + 
  facet_grid(~ set) + 
  theme_bw()
```

@@

## Assessing the Variability of the LIME Results

```{r}
# Perform the sensitivity analysis if not already saved
if(!file.exists("../data/hamby224_sensitivity_inputs.rds")) {

  # Specify the number of reps and input cases
  nreps = 10
  noptions = 12

  # Specify the inputs for the sensitivity analysis
  hamby224_sensitivity_inputs <- data.frame(bin_continuous = c(rep(TRUE, nreps * 10), 
                                                      rep(FALSE, nreps * 2)),
                                   quantile_bins = c(rep(c(TRUE, FALSE), each = nreps * 5), 
                                                     rep(TRUE, nreps * 2)),
                                   nbins = c(rep(rep(2:6, each = nreps), 2), 
                                             rep(4, nreps * 2)),
                                   use_density = c(rep(TRUE, nreps * 11), 
                                                   rep(FALSE, nreps))) %>%
    mutate(case = 1:(nreps * noptions)) %>%
    select(case, bin_continuous, nbins, quantile_bins, use_density)

  # Tell R to run the upcoming code in parallel
  plan(multiprocess)

  # Run lime for the sensitivity analysis
  hamby224_sensitivity_outputs <- future_pmap(.l = as.list(hamby224_sensitivity_inputs %>%
                                                             select(-case)),
             .f = run_lime, # run_lime is one of my helper functions
             train = hamby173and252_train %>% select(rf_features),
             test = hamby224_test %>% arrange(case) %>% select(rf_features) %>% na.omit(),
             rfmodel = as_classifier(rtrees),
             label = "TRUE",
             nfeatures = 3,
             seed = FALSE) %>%
    map_df(function(list) list$explain) %>%
    mutate(rep = factor(rep(rep(1:nreps, each = dim(hamby224_test %>% na.omit())[1] * 3),
                            noptions))) %>%
    full_join(hamby224_test %>% na.omit() %>% 
                mutate(case = as.character(case)), by = "case") %>%
    mutate(case = factor(case)) %>%
    select(case, model_r2:feature_weight, bin_continuous:rep, set:land2, rfscore, samesource)

  # Save the sensitivity inputs and outputs
  saveRDS(hamby224_sensitivity_inputs, "../data/hamby224_sensitivity_inputs.rds")
  saveRDS(hamby224_sensitivity_outputs, "../data/hamby224_sensitivity_outputs.rds")

  # Load in the sensitivity inputs and outputs
  hamby224_sensitivity_inputs <- readRDS("../data/hamby224_sensitivity_inputs.rds")
  hamby224_sensitivity_outputs <- readRDS("../data/hamby224_sensitivity_outputs.rds")

```{r}
hamby224_sensitivity_results <- hamby224_sensitivity_outputs %>%
  mutate(situation = ifelse(bin_continuous == TRUE & quantile_bins == TRUE, 
                            sprintf("%.0f quantile bins", nbins),
                            ifelse(bin_continuous == TRUE & quantile_bins == FALSE, 
                                   sprintf("%.0f equally space bins", nbins),
                                   ifelse(bin_continuous == FALSE & use_density == TRUE, 
                                          "kernel density", "normal approximation"))),
         diff = rfscore - model_prediction) %>%
  mutate(situation = fct_relevel(situation, 
                                 "2 quantile bins", 
                                 "3 quantile bins",
                                 "4 quantile bins",
                                 "5 quantile bins",
                                 "6 quantile bins"))

hamby224_sensitivity_mses <- hamby224_sensitivity_results %>%
  group_by(set, rep, situation) %>%
  summarise(mse = sum((diff)^2) / length(diff))
```

```{r}
hamby224_sensitivity_results %>%
  group_by(set, case, rep, situation) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(set, case, situation) %>%
  summarise(nlevels = length(levels(factor(feature))),
            count = n()) %>%
  ungroup() %>%
  group_by(set, situation) %>%
  summarise(mean_nlevels = mean(nlevels),
            sd_nlevels = sd(nlevels),
            upper = mean(nlevels) + sd(nlevels),
            lower = mean(nlevels) - sd(nlevels)) %>%
  ggplot(aes(x = situation, y = mean_nlevels)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = lower, ymax = upper)) +
  facet_grid(. ~ set) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
dat <- data.frame()

hamby224_sensitivity_results %>%
  filter(set == "Set 1") %>%
  group_by(case, situation, rep) %>%
  slice(1) %>%
  ggplot(aes(x = case, fill = feature)) + 
  geom_bar(position = "stack") + 
  coord_flip() + 
  facet_wrap(situation ~ ., ncol = 5) + 
  theme(legend.position = "bottom")
```

```{r}
hamby224_sensitivity_mses %>%
  group_by(set, situation) %>%
  summarise(mean = mean(mse),
            sd = sd(mse))
```

```{r}
hamby224_sensitivity_mses %>%
  ggplot(aes(x = situation, y = mse)) + 
  geom_boxplot() + 
  geom_point() + 
  facet_grid( ~ set) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "", y = "MSE")
```

@@

Ideas for improvement of LIME: 
  - change the binning method:
    - lime is not good with linear relationships with classifiers due to the 
    inside-outside binning - it would be better to use a cumulative approach
  - consider interactions in the ridge regression model
  - learn lambda in the ridge regression
  - could add a penalty for the number of parameters in the model (or the number of bins)
  - could start with a fine grid of bins and then go backwards and fit the models on this sampled data
  - could try a type of ANOVA test if we can have nested models or if we can assume nesting
  - could compute something like an R^2_adj
  - could force the intercept to be 0.5
  - try out the subsampling
